Malacology: A Programmable Storage System

### 2016-10-21 paper/paper.tex
paper: transitions for mantle eval
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Our evaluation demonstrates the feasibility of building new service abstractions
atop programmable storage, focusing on the
performance of the internal
abstractions exposed by Malacology and
used to construct the Mantle and ZLog services. We also discuss
latent capabilities we discovered in this process that allow us---as well as future
storage system programmers---to navigate
different trade-offs within the services themselves. First, we benchmark
scenarios 
%where contention is at the sequencer 
with high sequencer contention
by examining the interfaces used
to map ZLog onto Malacology; specifically, we describe the sequencer
implementation and the propagation of object and data interfaces interfaces.
Next, we benchmark scenarios in which the storage system manages multiple logs by using Mantle to
balance sequencers across a cluster.

@@

Since this work
focuses on the programmabilty of Malacology, the goal of this section is to
show that the components and subsystems that support the Malacology interfaces
provide reasonable relative performance, as well as to give examples of the flexibility
that Malacology provides to programmers.
%As we search for the best metadata balancing policies
%and physical design implementation trade-offs,
%we plan to rel

@@

For scenarios with multiple shared logs and many clients, co-locating
\paa{you might want a sentence or two of high-level explanation of this use case.
in practice, a storage system implementing CORFU will support a multiplicity of independent
totally-ordered logs (eg, one for each application).}
sequencers on the same physical node is not ideal. Building a
load balancer that can migrate the shared resource (in this case, the resource
that mediates access to the tail of the log) is a time-consuming, non-trivial task that requires building subsystems
for migrating resources, monitoring the workloads, collecting metrics that
describe the utilization on the physical nodes, partitioning resources,
maintaining cache coherence, and managing multiple sequencers. The following
experiments demonstrate the feasibility of using the mechanisms of the
Malacology load balancing interface to inherit all these features and to
alleviate load from overloaded servers with multiple sequencers.

@@

Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput (\(y\) axis)
over time (\(x\) axis) of different load balancers that migrate ZLog sequencers
acround the cluster; ``No Balancing" keeps all sequencers on one server,
``CephFS" migrates sequencers using the original algorithms packaged with Ceph,
and ``Mantle" uses a custom load balancer we wrote specifically for
sequencers. The increased throughput for the CephFS and Mantle curves between 0
and 60 seconds are a result of migrating the sequencer(s) off overloadeed
servers. In addition to showing that migrating sequencers improves performance,
Figure~\ref{fig:mantle-balancer-behaviors} also demonstrates features that we
will explore in the rest of this section.
Sections~\ref{sec:feature-balancing-modes}
and~\ref{sec:feature-migration-units} quantify the differences in performance
when the cluster stabilizes at time 100 seconds and
Section~\ref{sec:feature-backoff} examines the slope and start time of the
rebalancing phase between 0 and 60 seconds by comparing the aggressiveness of
the balancers.

@@

The load balancing service can operate in different balancing modes that stress
the subsystems that receive and forward client requests in different ways. In
Figure~\ref{fig:mantle-balancer-behaviors}, the CephFS curve shows the
performance of the balancing mode that CephFS uses {\it most of the time}.
CephFS has 3 modes for balancing load: CPU mode, workload mode, and hybrid
mode. All three have the same structure for making migration decisions but vary
based on the metric used to calculate load. For this sequencer workload, the 3
different modes all have the same performance, as shown in
Figure~\ref{fig:mantle-balancer-performance}, because the load balancer falls
into the same mode a majority of the time.  The high variation in performance
for the CephFS CPU Mode bar reflects the uncertainty of using something as
dynamic and unpredictable as CPU utilization to make migration decisions. In
addition to the suboptimal performance and unpredictability, the fact that the
CephFS balancers all fall into the same mode despite using different balancers
is a problem becuase it prevents administrators from properly exploring the
balancing state space.

@@

Figure~\ref{fig:mantle-mode-behavior} shows the throughput (\(y\) axis) over
time (\(x\) axis) of the two different modes for an environment with only 2
sequencers (again 4 clients each) and 2 servers. The curves for both sequencers
in Figure~\ref{fig:mantle-mode-behavior}(a) start at less than 1000 ops/second
and at time 60 seconds Mantle migrates Sequencer 1 to the slave server.
Performance of Sequencer 2 decreases because it stayed on the proxy which now
handles all client requests, processes requests for Sequencer 2, and forwards
requests for Sequencer 1. The performance of Sequencer 1 improves dramatically
because lookups for that sequencer are parallelized with the operations on the
other server. Cluster throughput improves at the cost of decreased throughput
for Sequencer 2.  Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer
mode manually (no balancing phase) and shows that the cluster throughput is
worse than the cluster throughput of proxy mode. That graph also shows that
Sequencer 2 has less throughput than Sequencer 2. In this case, the
scatter-gather process used for cache coherence in the metadata protocols
causes strain on the server housing Sequencer 2 resulting in this uneven
performance.

Intuitively, client mode is a superior balancing mode and should perform
better; after all, this is the best strategy for distributing POSIX metadata in
write-heavy workloads~\cite{ren:sc2014-indexfs, patil:fast2011-giga,
sevilla:sc15-mantle}.  But our experiments show that using proxy mode improves
performance for this smale scale setup because it separates (1) the handling of
the client requests and (2) finding the tail of the log and responding to
clients.  Doing both steps is too heavy weight for one server and sequencers on
slave nodes can go faster because work is split up; this phenomenon is not
uncommon and has been observed in chain replication~\cite{CITEME}.

@@

We define a ``migration unit" as the amount of load to try and migrate.  For
example, it is useful to express migration units as a fraction of the total
load.  Expressing this heuristic is not easily achievable with outward facing
tunable parameters (i.e. system knobs) and so the original CephFS balancers do
not have a migration unit tunable.  With programmability, migration units can
be expressed quite easily. To force the load balancer into the Proxy Mode (Half)
scenario in Figure~\ref{fig:mantle-modes}, which sets the balancer to proxy
mode with migration units of half the load on the current server we can use the
Mantle policy:\\

@@

The balancing phase from 0 to 60 seconds in
Figure~\ref{fig:mantle-balancer-behaviors} shows different degrees of
aggressiveness in making migration decisions; CephFS makes a decision quickly
and throughput jumps to 2500 ops/second within 10 seconds while Mantle takes
more time to stabilize. This conservative behavior is controlled by programming
the balancer to (1) use different conditions for when to migrate and (2) using
a threshold for sustained overload.

@@

%We use a balancer policy specifically designed for multiple sequencers. The
%balancer aggressively sheds half its load to the next server as soon as it can.
%It is based off the ``Greedy Spill Even" approach in the
%Mantle~\cite{sevilla:sc15-mantle} paper but is reproduced in
%Figure~\ref{listing:greedy-spill} for the reader's convenience. Each line of
%code uses variables exposed by the Mantle API and is executed as a callback by
%Mantle.  ``Server load" lets the administrator control what metric constitutes
%load on a server; ``i" is a variable used by Mantle to iterate over servers and
%``req\_rate" is a metric based on the exponenential decay of the request rate.
%``When policy" is used by Mantle to set a conditional for kicking off
%migration; in this case, the cluster migrates load if the current server,
%``whoami", has load (\texttt{>0.1}) and the next server (physical servers have
%``ranks") has a small amount load (\texttt{<0.1}). Finally, the ``where policy"
%is used by Mantle to determine how much load to shed. The policy fills in a
%``targets" array to tell the migrator how much load to send to each server; in
%this case, we shed half of the current load to the next server. The unit of
%load is a single sequencer for a ZLog log.
%
%\begin{figure}[]
%\begin{lstlisting}
%  -- Server load
%  load = server[i]["req_rate"]
%  
%  -- When policy
%  if server[whoami]["load"]>.01 and 
%    server[whoami+1]["load"]<.01 then
%    
%  -- Where policy
%  targets[whoami+1]=load/2
%\end{lstlisting}
%\caption{The sequencer balancing policy that works best for bursty workloads. 
%\label{listing:greedy-spill}}
%\end{figure}
%
%Figure~\ref{fig:mantle-seq-thruput} shows a performance profile of one of the
%sequencers as the load balancing interface migrates sequencers around the
%cluster to alleviate load.  Each curve shows the throughput (\(y\) axis) over
%time (\(x\) axis).  ``Underloaded (Baseline)'' is the performance of one
%sequencer on one physical server serving 7 clients - there are no other
%sequencers or logs in the system and the curve represents the maximum
%throughput of a sequencer on an underutilized physical server.  ``Overloaded:
%No Balancing" adds a second sequencer of equal load to the same physical server
%with the intent of overloading the single physical server; recall that the
%throughput for this second sequencer is omitted. Compared to the baseline, the
%overloaded physical server with 2 sequencers has more than a 2\(\times\)
%throughput degradation.  The stability of throughput curves (i.e. the variation
%on the \(y\) axis) indicate which sequencers are on overloaded physical
%servers.  ``Underloaded (Baseline)" is the best performance for this particular
%workload because there is no contention.  Performance is stable and the MDS is
%under-loaded because we do not see the spike at the end like the other curves.
%``Overloaded: No Balancing" shows an overloaded sequencer because the
%performance of the monitored sequencer is low and unstable (high variation in
%performance). 
%
%The last two curves of Figure~\ref{fig:mantle-seq-thruput} show the performance
%improvements when sequencers are load balanced across the cluster.  The load
%balancing policies differ by the threshold at which they start balancing:
%``Overloaded: Sequencer Policy" starts balancing immediately and ``Overloaded:
%Conservative Policy" waits for 120 seconds before migrating sequencers.
%
%%Overall, the clients achieve lower performance than the capability experiments
%%#in Figure~\ref{fig:captp} because the sequencer uses a slower metadata
%%operation to manage the capability.  It appears the bottleneck is somewhere
%%with the clients -- perhaps it is our 1GBit/s network link.
%
%Figure~\ref{fig:mantle-cpu-utilization} shows how load balancing effects the
%physical servers by plotting the CPU utilization average over 1 minute (\(y\)
%axis) over time (\(x\) axis).  It shows the ``Conservative Policy" load
%balancer from the run in Figure~\ref{fig:mantle-seq-thruput}, so recall that
%resources are constrained during the job by co-locating a second sequencer on
%the same physical server.  CPU utilization spikes when physical servers receive
%load (at 2 minutes physical server 1 receives a sequencer from physical server
%0) but the CPU utilization never eclipses 50\%. The utilization is low because
%the load balancer interface is not multi-threaded.  Recall the the load
%balancing interface re-uses the POSIX metadata cluster migration mechanisms;
%serving file system metadata is single-threaded because it has been shown that
%metadata performance does not improve with multiple
%threads~\cite{konstantinos:pdsw2014-lustre-metadata}.

\noindent\textbf{Malacology pulls} the load balancing out of the storage system
to balance sequencers across a cluster. Compared to putting all sequencers on
the same server, load balancing a sequencer using the proxy mode shows a
1.8\(\times\) throughput improvement.\\

\noindent\textbf{Features:} different load balancing modes control forwarding
vs. client redirection, tuning the load migration unit parameter controls
sequencer distribution vs. co-location, and tuning the backoff parameter
controls conservative vs.  aggressive decision making.

>>>>

Our evaluation demonstrates the feasibility of building new service
abstractions atop programmable storage, focusing on the performance of the
internal abstractions exposed by Malacology and used to construct the Mantle
and ZLog services. We also discuss latent capabilities we discovered in this
process that let us navigate different trade-offs within the services
themselves. First, we benchmark scenarios with high sequencer contention by
examining the interfaces used to map ZLog onto Malacology; specifically, we
describe the sequencer implementation and the propagation of object and data
interfaces interfaces.  Next, we benchmark scenarios in which the storage
system manages multiple logs by using Mantle to balance sequencers across a
cluster.

@@

Since this work focuses on the programmabilty of Malacology, the goal of this
section is to show that the components and subsystems that support the
Malacology interfaces provide reasonable relative performance, as well as to
give examples of the flexibility that Malacology provides to programmers.

%As we search for the best metadata balancing policies
%and physical design implementation trade-offs,
%we plan to rel

@@

In practice, a storage system implementing CORFU will support a multiplicity of
independent totally-ordered logs for each application.  For this scenario
co-locating sequencers on the same physical node is not ideal. Building a load
balancer that can migrate the shared resource (in this case, the resource that
mediates access to the tail of the log) is a time-consuming, non-trivial task
that requires building subsystems for migrating resources, monitoring the
workloads, collecting metrics that describe the utilization on the physical
nodes, partitioning resources, maintaining cache coherence, and managing
multiple sequencers. The following experiments demonstrate the feasibility of
using the mechanisms of the Malacology load balancing interface to inherit all
these features and to alleviate load from overloaded servers housing multiple
sequencers.

@@

First, we show how the ZLog service can orchestrate multiple sequencers in a
cluster using the Malacology load balancing interface.
Figure~\ref{fig:mantle-balancer-behaviors} shows the throughput (\(y\) axis)
over time (\(x\) axis) of different load balancers that migrate ZLog sequencers
acround the cluster; ``No Balancing" keeps all sequencers on one server,
``CephFS" migrates sequencers using the CephFS load balancers, and ``Mantle"
uses a custom load balancer we wrote specifically for sequencers. The CephFS
load balancers are hard-coded into Ceph while Mantle is an API for
balancing load presented in~Mantle~\cite{sevilla:sc15-mantle} and
re-implemented on top of Malacology in this paper.  The increased throughput
for the CephFS and Mantle curves between 0 and 60 seconds are a result of
migrating the sequencer(s) off overloadeed servers.

In addition to showing that migrating sequencers improves performance,
Figure~\ref{fig:mantle-balancer-behaviors} also demonstrates features that we
will explore in the rest of this section.
Sections~\ref{sec:feature-balancing-modes}
and~\ref{sec:feature-migration-units} quantify the differences in performance
when the cluster stabilizes at time 100 seconds and
Section~\ref{sec:feature-backoff} examines the slope and start time of the
rebalancing phase between 0 and 60 seconds by comparing the aggressiveness of
the balancers.

@@

Next, we quantify the performance benefits shown in
Figure~\ref{fig:mantle-balancer-behaviors}.  To understand why load balancers
perform differently we need to explain the different balancing modes that the
load balancer service uses and how they stress the subsystems that receive and
forward client requests in different ways. In
Figure~\ref{fig:mantle-balancer-behaviors}, the CephFS curve shows the
performance of the balancing mode that CephFS falls into {\it most of the time}.
CephFS has 3 modes for balancing load: CPU mode, workload mode, and hybrid
mode. All three have the same structure for making migration decisions but vary
based on the metric used to calculate load. For this sequencer workload the 3
different modes all have the same performance, shown in
Figure~\ref{fig:mantle-balancer-performance}, because the load balancer falls
into the same mode a majority of the time.  The high variation in performance
for the CephFS CPU Mode bar reflects the uncertainty of using something as
dynamic and unpredictable as CPU utilization to make migration decisions. In
addition to the suboptimal performance and unpredictability, the fact that the
CephFS balancers all fall into the same mode despite using different balancers
is a problem becuase it prevents administrators from properly exploring the
balancing state space.

@@

Figure~\ref{fig:mantle-mode-behavior} shows the throughput over time of the two
different modes for an environment with only 2 sequencers (again 4 clients
each) and 2 servers. The curves for both sequencers in
Figure~\ref{fig:mantle-mode-behavior}(a) start at less than 1000 ops/second and
at time 60 seconds Mantle migrates Sequencer 1 to the slave server.
Performance of Sequencer 2 decreases because it stayed on the proxy which now
handles all client requests, processes requests for Sequencer 2, and forwards
requests for Sequencer 1. The performance of Sequencer 1 improves dramatically
because distributing the sequencers in this way separates (1) the handling of
the client requests and (2) finding the tail of the log and responding to
clients.  Doing both steps is too heavy weight for one server and sequencers on
slave nodes can go faster because work is split up; this phenomenon is not
uncommon and has been observed in chain replication~\cite{CITEME}.  Cluster
throughput improves at the cost of decreased throughput for Sequencer 2.
Figure~\ref{fig:mantle-mode-behavior}(b) is set to sequencer mode manually (no
balancing phase) and shows that the cluster throughput is worse than the
cluster throughput of proxy mode. That graph also shows that Sequencer 2 has
less throughput than Sequencer 2. In this case, the scatter-gather process used
for cache coherence in the metadata protocols causes strain on the server
housing Sequencer 2 resulting in this uneven performance.

@@

Another factor that affects performance in this environment is ``how much" load
is on each server; these experiments quantify that effect by programming
policies into the load balancing interface to control the amount of load to
migrate. We defined a  migration unit as the amount of load to try and migrate.
For example, it is useful to express migration units as a fraction of the total
load.  Expressing this heuristic is not easily achievable with outward facing
tunable parameters (i.e. system knobs) and so the original CephFS balancers do
not have a migration unit tunable.  With programmability, migration units can
be expressed quite easily. To force the load balancer into the Proxy Mode
(Half) scenario in Figure~\ref{fig:mantle-modes}, which sets the balancer to
proxy mode with migration units of half the load on the current server we can
use the Mantle policy:\\

@@

Tuning the aggressiveness of the load balancer decision making is also a
trade-off that administrators can control and explore. The balancing phase from
0 to 60 seconds in Figure~\ref{fig:mantle-balancer-behaviors} shows different
degrees of aggressiveness in making migration decisions; CephFS makes a
decision quickly and throughput jumps to 2500 ops/second within 10 seconds
while Mantle takes more time to stabilize. This conservative behavior is
controlled by programming the balancer to (1) use different conditions for when
to migrate and (2) using a threshold for sustained overload.

@@

\noindent\textbf{Malacology pulls} the load balancing out of the storage system
to balance sequencers across a cluster. Compared to putting all sequencers on
the same server, load balancing a sequencer using the proxy mode shows a
1.8\(\times\) throughput improvement. This latent capability also gives future
programmers the ability to explore the different load balancing trade-offs
including: load balancing modes to control forwarding vs. client redirection,
load migration units to control sequencer distribution vs. co-location, and
backoffs to control conservative vs. aggressive decision making.
